%!TEX root = ../notes.tex
\section{March 30, 2026}
\label{20260330}

\subsection{Multi-Party Computation - Big Picture}

There are two different levels of MPC: either semi-honest or malicious. Semi-honest means that the adversary follows the protocol, but tries to extract more information. Malicious means that the adversary deviates from the protocol to extract information.

If we have an oblivious transfer (OT) that is secure against semi-honest adversaries, we can use it with Yao's garbled circuit to achieve semi-honest 2-party computation for any function. Then, with the cut-and-choose with commitments, we can achieve malicious 2-party computation. We will discuss this next.

\begin{align*}
    \text{Semi-honest OT}\ & \underset{\text{Yao's Garbled Circuits}}{\implies}\ \text{Semi-honest 2PC for any function} \\
    & \underset{\text{Cut-and-choose with commitments}}{\implies}\ \text{Malicious 2PC for any function}
\end{align*}

In the last part of this lecture, we will discuss how to use semi-honest OT with GMW to achieve semi-honest multi-party computation, which can be extended to malicious MPC.

\begin{align*}
    \text{Semi-honest OT} \ & \underset{\text{GMW}}{\implies} \text{Semi-honest MPC for any function}\\
    & \underset{\text{GMW Compiler with ZKP}}{\implies} \text{Malicious MPC for any function}
\end{align*}

Recall the definition of Oblivious Transfer.

\begin{definition}[Oblivious Transfer]
    An \ul{oblivious transfer} is a protocol in which a sender, with messages $m_0, m_1\in\{0, 1\}^l$ gives a choice to the receiver to receive either $m_0, m_1$.

    Given a choice bit from the receiver $b\in\{0,1\}$, the receiver gets $m_b$ and the sender also gets no information about the message transferred.

    % \Graphic{images/2023-03-21/ot.png}{0.6}

    \pseudocodeblock{
        \textbf{Sender} \< \< \textbf{Receiver}\\
        \text{Input: } m_0, m_1 \in \{ 0 , 1\}^{\ell} \< \< \text{Input: } b \in \{0, 1\}\\
        \< \sendmessageright*{ } \< \\
        \< \sendmessageleft*{ }\< \\
        \< \sendmessageright*{ }\< \\
        \< \sendmessageleft*{ } \< \\
        \text{Output: }\perp \< \< \text{Output: }m_b
    }
\end{definition}
